Bridging the Gap Between Visual and Analytical Machine Learning Testing

### 2023-10-25 report/icst24.tex
typo indent
adjust implications again

@@

@@

Additionally, in the results from RQ2\luis{this is RQ3}, we see several examples of assertions that fail to capture all the information gained from its visual counter-part. We argue that this can lead to a mismatch of expectations in future versions of the ML software: for example, another collaborator of the project might run the code with a new version of the dataset or a different learning technique and assume that if the assertions run successfully, the expected patterns from the visualisation are also being met. After delving into some examples, we also observe that deriving assertions from validations is far from trivial: basic visualisation patterns require complex mathematical techniques that might be out of the expertise of the ML developer. We ought to bridge the gap between qualitative assessment of ML systems using visualisations and deriving analytical assertions from the visualisations. In other words, there is a need for automated tools that make this translation process easier for AI practitioners. In sum, our results showcase a gap in the current state of the art in software testing: it is failing to address the idiosyncrasies of ML code.

>>>>

\luis{Amother thought: there are still many visualisations without asserts. We believe in each visualisation there is an expectation that has not been formalised and that it will not be easily validated by other contributors. this is already in the motivation but we can frame it as a visible open problem here as well.}

@@

\luis{some thought that is blended here but maybe we want to make it clear. ML testing is a task that requires practitioners to have a strong background in machine learning, software testing, and mathematics (or statistics). This reinforces the need to adjust the education of these professionals and to design tools that aim mitigate the lack of background in some of these directions.}

@@

Additionally, in the results from RQ2\luis{this is RQ3}, we see several examples of assertions that fail to capture all the information gained from its visual counter-part. We argue that this can lead to a mismatch of expectations in future versions of the ML software: for example, another collaborator of the project might run the code with a new version of the dataset or a different learning technique and assume that if the assertions run successfully, the expected patterns from the visualisation are also being met. After delving into some examples, we also observe that deriving assertions from validations is far from trivial: basic visualisation patterns require complex mathematical techniques that might be out of the expertise of the ML developer. For example, the linear correlation observed with the scatter plot of Figure 6\luis{add ref} is clear to any non-expert but it had to be formalised (cf. Listing 11\luis{add ref}) using the metric coefficient of determination bounded to a sound threshold (0.6). Hence, we ought to bridge the gap between qualitative assessment of ML systems using visualisations and deriving analytical assertions from the visualisations. In other words, there is a need for automated tools that make this translation process easier for AI practitioners. In sum, our results showcase a gap in the current state of the art in software testing: it is failing to address the idiosyncrasies of ML code.
